Recursive Solutions to First-Order Model Counting

### 2022-02-17 notes/notes.tex
Some small changes

\newcommand\pfun{\mathrel{\ooalign{\hfil$\mapstochar\mkern5mu$\hfil\cr$\to$\cr}}}

@@

\paragraph{Things I might need to explain.}
\begin{itemize}
\item atom, (positive/negative) literal, constant, predicate, variable, literal variable, clause, unit clause
\item $\Vars$, $\Vars(c) = \Vars(L) \cup \Vars(C)$
\item $\Doms$ on both formulas and clauses. $\Doms(c) = \Imm \delta_c$, and $\Doms(\phi) = \bigcup_{c \in \phi} \Doms(c)$.
\item maybe: notation for partial function, notation for powerset, notation for disjoint union, domain size
\item $\WMC$, $\wwp$, $\wwn$, $\Imm$
\item two parts: compilation and inference.
\item introduce and use arrows for bijections, injections, set inclusions, etc.
\item $\emptyset$ for an empty partial map.
\item notation for variable substitution for literal and constraint sets: $S[x/V]$, where $S$ is the set, $V$ is a set of variable, and $x$ is the new constant or variable. Note that for constraint sets this could mean that a variable and a constant need to switch places.
\item During inference, there is a domain size map $\size\colon \mathscr{D} \to \mathbb{N}_0$.
\item $\sqcup$ for both sets and functions
\item mention which rules are in $\Gamma$ and which ones are in $\Delta$ (and why tryCache has to be in $\Delta$).
\end{itemize}

\paragraph{TODO}
\begin{itemize}
\item should I always say `compilation rule' instead of `operator'?
\item maybe mathcal instead of mathscr
\item maybe $\pi$ is global enough to have a more unique name
\item maybe $c[\beta, \gamma]$ is better than $c \ast (\beta, \gamma)$
\item capitalise variable names.
\item introduce/use notation for vertices, edges of a graph
\item maybe `plus-plus' is better than $\oplus$
\item maybe put each function into its own algorithm environment, using the caption for the function name
\end{itemize}

@@

\begin{definition}
  A \emph{domain} is a symbol for a finite set.\footnote{In the context of functions, the domain of a function $f$ retains its usual meaning and is denoted $\dom(f)$.} Let $\mathscr{D}$ be the set of all domains and $\mathscr{C} \subset \mathscr{D}$ be the subset of domains introduced as a consequence of constraint removal. Note that both sets (can) expand during the compilation phase.

Let $\pi\colon \mathscr{D} \pfun \mathscr{D}$ be a partial endomorphism on $\mathscr{D}$ that denotes the \emph{parent} relation, i.e., if $\pi(d) = e$ for some $d, e \in \mathscr{D}$, then we call $e$ the parent (domain) of $d$, and $e$ a child of $d$. Intuitively, $\pi$ arranges all domains into a forest---thus, we often use graph theoretical terminology to describe properties of and relationships between domains.
\end{definition}

@@

% TODO: so far I'm not using equivalence anywhere
\begin{definition}
  A \emph{clause} is a triple $c = (L, C, \delta_c)$, where $L$ is the set of literals, $C$ is a set of inequality constraints, and $\delta_c\colon \Vars(c) \to \mathscr{D}$ is a function that maps all variables in $c$ to their domains such that if $(x, y) \in C$ for some variables $x$ and $y$, then $\delta_c(x) = \delta_c(y)$.

  Two clauses $c$ and $d = (L', C', \delta_d)$ are \emph{isomorphic} (written $c \cong d$) if there is a bijection $\beta\colon \Vars(c) \to \Vars(d)$ such that $c\beta = d\beta$. TODO: we will always use this subscript notation for the $\delta$'s. Equality of clauses is defined in the usual way (i.e., all variables, domains, etc. must match).
\end{definition}

@@

@@

\begin{definition}
  A \emph{first-order deterministic decomposable negation normal form computational graph} (FCG) is a (weakly connected) directed graph with a a single source, node labels, and ordered direct successors. We denote an FCG as $G = (V, s, N^+, \lambda)$, where $V$ is the set of nodes, $s \in V$ is the source, $N^+$ is the direct successor function that maps each node in $V$ to a list that contains either other nodes in $V$ or $\star$, which means that the target of the edge is yet to be determined, and $\lambda\colon V \to \mathscr{L}$ is the vertex-labelling function with some codomain $\mathscr{L}$, which is assumed to be the same for all FCGs. For each $v \in V$, the length of the list $N^+(v)$ is determined by $\lambda(v)$.

@@

\paragraph{TODO.}
\begin{itemize}
\item explain the connection to the algebraic formalism (or just don't use it). % We will also use algebraic notation to describe (parts of) circuits: circuittype_parameters(nodes pointed to by the outedges)
\item define `an FCG for a formula'
\item vertex or node, $v$ or $n$?
\item what notation for lists (a.k.a. sequences) am I going to use? $\oplus$ for concatenation (of both two lists and an element to a list), $\in$ for (in-order) enumeration, $[]$ for an empty list, $[x]$ for a list with one element, $|L|$ for the length of the list, and $h : t$ to denote a list with first element (a.k.a. head) $h$ and remaining list (a.k.a. tail) $t$. Also add an example of the list comprehension notation.
\item use Greek letters for compilation rules
\end{itemize}

@@

\begin{algorithm}[t]
  \caption{The (main part of the) search algorithm}
  \label{alg:search}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKwData{solutions}{solutions}
  \SetKwFunction{applyGreedyRules}{applyGreedyRules}
  \SetKwFunction{applyNonGreedyRules}{applyNonGreedyRules}
  \SetKwFunction{put}{put}
  \SetKwFunction{get}{get}
  \SetKwFunction{emptyy}{empty}
  \KwIn{a formula $\phi_0$}
  \KwResult{all found FCGs for $\phi_0$ are in the set \solutions}
  $\solutions \gets \emptyset$\;
  $C_0 \gets \emptyset$\;
  $(G_0, C_0, L_0) \gets \applyGreedyRules{$\phi_0$, $C_0$}$\;
  \lIf{$L_0 = []$}{$\solutions \gets \{\, G_0 \,\}$}
  \Else{
    $q \gets \text{an empty queue of states}$\;
    $q.\put{$(G_0, C_0, L_0)$}$\;
    \While{{\bf not} $q.\emptyy{}$}{
      \ForEach{state $(G, C, L) \in \applyNonGreedyRules{$q.\get{}$}$}{
        \lIf{$L = []$}{$\solutions \gets \solutions \cup \{\, G \,\}$}
        \lElse{$q.\put{$(G, C, L)$}$}
      }
    }
  }
\end{algorithm}

\begin{algorithm}
  \caption{Functions used in \cref{alg:search} for applying compilation rules}
  \label{alg:apply}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKwData{newStates}{newStates}
  \SetKwFunction{applyGreedyRules}{applyGreedyRules}
  \SetKwFunction{applyNonGreedyRules}{applyNonGreedyRules}
  \SetKwFunction{applyGreedyRulesToAllFormulas}{applyGreedyRulesToAllFormulas}
  \SetKwFunction{mergeFcgs}{mergeFcgs}
  \SetKwFunction{updateCache}{updateCache}
  \KwData{a set of greedy rules $\Gamma$}
  \KwData{a set of non-greedy rules $\Delta$}
  \Fn{\applyGreedyRules{$\phi$, $C$}}{
    \ForEach{rule $r \in \Gamma$}{
      $S \gets r(\phi)$\;
      \If{$S \ne \emptyset$}{
        $(G, L) \gets \text{any element of } S$\;
        \lIf{$G = {\normalfont \texttt{null}}$}{\Return{\applyGreedyRules{the only element of $L$, $C$}}}
        \Else{
          $(V, s, N^+, \lambda) \gets G$\;
          $C \gets \updateCache(C, \phi, s)$\;
          \Return{\applyGreedyRulesToAllFormulas{$G$, $C$, $L$}}\;
        }
      }
    }
    \Return{$({\normalfont \texttt{null}}, C, [\phi])$}\;
  }
  \Fn{\applyGreedyRulesToAllFormulas{$G$, $C$, $L$}}{
    $(V, s, N^+, \lambda) \gets G$\;
    $N^+(s) \gets []$\;
    $L' \gets []$\;
    \ForEach{formula $\phi \in L$}{
      $(G', C, L'') \gets \applyGreedyRules{$\phi$, $C$}$\;
      $L' \gets L' \oplus L''$\;
      \lIf{$G' = {\normalfont \texttt{null}}$}{$N^+(s) \gets N^+(s) \oplus [\star]$}
      \Else{
        $(V', s', N', \lambda') \gets G'$\;
        $V \gets V \sqcup V'$\;
        $N^+ \gets N^+ \sqcup N'$\;
        $N^+(s) \gets N^+(s) \oplus [s']$\;
        $\lambda \gets \lambda \sqcup \lambda'$\;
      }
    }
    \Return{$((V, s, N^+, \lambda), C, L')$}\;
  }
  \Fn{\applyNonGreedyRules{$s$}}{
    $(G, C, L) \gets s$\;
    $\phi : T \gets L$\;
    $(G', C', L') \gets \text{a copy of } s$\;
    $\newStates \gets []$\;
    \ForEach{rule $r \in \Delta$}{
      \ForEach{$(G'', L'') \in r(\phi)$}{
        \lIf{$G'' = {\normalfont \texttt{null}}$}{$\newStates \gets \newStates \oplus \applyNonGreedyRules{$(G', C', L'')$}$}
        \Else{
          $(V, s, N^+, \lambda) \gets G''$\;
          $C' \gets \updateCache{$C'$, $\phi$, $s$}$\;
          $(G'', C', L'') \gets \applyGreedyRulesToAllFormulas{$G''$, $C'$, $L''$}$\;
          \lIf{$G' = {\normalfont \texttt{null}}$}{$\newStates \gets \newStates \oplus [(G'', C', L'' \oplus T)]$}
          \lElse{$\newStates \gets \newStates \oplus [(\mergeFcgs{$G'$, $G''$}, C', L'' \oplus T)]$}
        }
      }
      $(G', C', L') \gets \text{a copy of } s$\;
    }
    \Return{\newStates}\;
  }
\end{algorithm}

\begin{algorithm}
  \caption{Helper functions used by \cref{alg:apply}}
  \label{alg:helpers}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKwFunction{mergeFcgs}{mergeFcgs}
  \SetKwFunction{updateCache}{updateCache}
  \Fn{\updateCache{$C$, $\phi$, $v$}}{
    \lIf{$\#\phi \not\in \dom(C)$}{\Return{$C \cup \{\, \#\phi \mapsto (\phi, v) \,\}$}}
    \lIf{there is no $(\phi', v') \in C(\#\phi)$ such that $v' = v$}{$C(\#\phi) \gets (\phi, v) \oplus C(\#\phi)$}
    \Return{$C$}\;
  }
  \Fn{\mergeFcgs{$G = (V, s, N^+, \lambda)$, $G' = (V', s', N', \lambda')$, $r = s$}}{
    \lIf{$\lambda(r) = \textsc{Ref}$}{\Return{\normalfont \texttt{null}}}
    \ForEach{$t \in N^+(r)$}{
      \If{$t = \star$}{
        replace $t$ with $s'$ in $N^+(r)$\;
        \Return{$(V \sqcup V', s, N^+ \sqcup N', \lambda \sqcup \lambda')$}\;
      }
      $G'' \gets \mergeFcgs{$G$, $G'$, $t$}$\;
      \lIf{$G'' \ne {\normalfont \texttt{null}}$}{\Return{$G''$}}
    }
    \Return{\normalfont \texttt{null}}\;
  }
\end{algorithm}

\paragraph{TODO}
\begin{itemize}
\item maybe `applyAllRules' is a better name
\item explain the `tail' part of the algorithm
\end{itemize}

@@

\begin{algorithm}
  \SetKwData{changed}{changed}
  \caption{Propagating atoms for smoothing across the FCG in a way that avoids infinite loops}
  \label{alg:smoothing}
  \KwIn{FCG $(V, s, N^+, \lambda)$}
  \KwIn{function $\iota$ that maps labels in $\mathscr{L}$ to sets of atoms}
  \KwIn{functions $\{\,f_l\,\}_{l \in \mathscr{L}}$ that take a list of sets of atoms and return a set of atoms}
  \KwOut{function $S$ that maps vertices in $V$ to sets of atoms}
  $S \gets \{\, v \mapsto \iota(\lambda(v)) \mid v \in V \,\}$\;
  $\changed \gets \texttt{true}$\;
  \While{\changed}{
    $\changed \gets \texttt{false}$\;
    \ForEach{vertex $v \in V$}{
      $S' \gets f_{\lambda(v)}([S(w) \mid w \in N^+(v)])$\;
      \If{$S' \ne S(v)$}{
        $\changed \gets \texttt{true}$\;
        $S(v) \gets S'$\;
      }
    }
  }
\end{algorithm}

@@

\begin{definition}[Notation]
  For any clause $c = (L, C, \delta_c)$, bijection $\beta\colon \Vars(c) \to V$ (for some set of variables $V$) and function $\gamma\colon \Doms(c) \to \mathscr{D}$, let $c \ast (\beta, \gamma) = d$ be the clause with all occurrences of any variable $v \in \Vars(c)$ in $L$ and $C$ replaced with $\beta(v)$ (so $\Vars(d) = V$) and $\delta_d\colon V \to \mathscr{D}$ defined as $\delta_d \coloneqq \gamma \circ \delta_c \circ \beta^{-1}$. In other words, $\delta_d$ is the unique function that makes the following diagram commute:
  \[
  \begin{tikzcd}
    \Vars(c) \ar[r, tail, two heads, "\beta"] \arrow[d, swap, "\delta_c"] & V = \Vars(d) \ar[d, dashed, "\exists!\delta_d"] \\
    \Doms(c) \ar[r, swap, "\gamma"] & \mathscr{D}.
  \end{tikzcd}
  \]
\end{definition}

\begin{algorithm}[t]
  \caption{A recursive function for checking whether one can reuse the circuit for computing $\WMC(\phi)$ to compute $\WMC(\psi)$. Both $\phi$ and $\psi$ are formulas, and $\rho\colon \Doms(\phi) \protect\pfun \Doms(\psi)$ is a partial map.}
  \label{alg:recursion}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKw{KwRet}{yield}
  \SetKwFunction{identifyRecursion}{identifyRecursion}
  \SetKwFunction{findHistory}{traceAncestors}
  \SetKwFunction{generateMaps}{generateMaps}
  \SetKwFunction{constructDomainMap}{constructDomainMap}
  \SetKwData{suitable}{suitableBijection}
  \SetKwData{found}{foundConstraintRemoval}
  \Fn{\identifyRecursion{$\phi$, $\psi$, $\rho = \emptyset$, $\found = {\normalfont \texttt{false}}$}}{
    \lIf{$|\phi| \ne |\psi|$ {\bf or} $\#\phi \ne \#\psi$}{\Return{\normalfont \texttt{null}}}
    \If{$\phi = \emptyset$}{
      \lIf{$\found$}{\Return{$\rho$}}
      \Return{\normalfont \texttt{null}}\;
    }
    \ForEach{clause $c \in \phi$\label{line:for1}}{
      \ForEach{clause $d \in \psi$ such that $\#d = \#c$\label{line:for2}} {
        \ForEach{$(\beta, \gamma) \in \generateMaps{$c$, $d$, $\rho$}$ such that $c \ast (\beta, \gamma) = d$\label{line:generateMaps}}{
          $\found' \gets \found$\;
          $\suitable \gets \texttt{true}$\;
          \ForEach{$e \in \Doms(c)$\label{line:vars}}{
            $\found'' \gets \findHistory{$e$, $\gamma(e)$}$\;
            \If{$\found'' = {\normalfont \texttt{null}}$}{
              $\suitable \gets \texttt{false}$\;
              break\;
            }
            \lIf{$\found''$}{$\found' \gets \texttt{true}$}
          }
          \If{\suitable}{
            $\rho'' \gets \identifyRecursion{$\phi \setminus \{\, c \,\}$, $\psi \setminus \{\, d \,\}$, $\rho \cup \gamma$, $\found'$}$\;\label{line:recursion}
            \lIf{$\rho'' \ne {\normalfont \texttt{null}}$}{\Return{$\rho''$}}
          }
        }
      }
      \Return{\normalfont \texttt{null}}\;
    }
  }
  \Fn{\generateMaps{$c$, $d$, $\rho$}}{
    \ForEach{bijection $\beta\colon \Vars(c) \to \Vars(d)$\label{line:bijection}}{
      $\gamma \gets \constructDomainMap{$\Vars(c)$, $\delta_c$, $\delta_d$, $\beta$, $\rho$}$\;
      \lIf{$\gamma \ne {\normalfont \texttt{null}}$}{\KwRet{$(\beta, \gamma)$}}
    }
  }
  \Fn{\constructDomainMap{$V$, $\delta_c$, $\delta_d$, $\beta$, $\rho$}}{
    $\gamma \gets \emptyset$\;
    \ForEach{$v \in V$}{
      \lIf{$\delta_c(v) \in \dom(\rho)$ {\bf and} $\rho(\delta_c(v)) \ne \delta_d(\beta(v))$}{\Return{\normalfont \texttt{null}}}
      \lIf{$\delta_c(v) \not\in \dom(\gamma)$}{$\gamma \gets \gamma \cup \{\, \delta_c(v) \mapsto \delta_d(\beta(v)) \,\}$}
      \lElseIf{$\gamma(\delta_c(v)) \ne \delta_d(\beta(v))$}{\Return{\normalfont \texttt{null}}}
    }
    \Return{$\gamma$}\;
  }
  \Fn{\findHistory{$d$, $e$}}{
    $\found \gets \texttt{false}$\;
    \While{$e \ne d$ {\bf and} $e \in {\normalfont \dom}(\pi)$}{
      \lIf{$e \in \mathscr{C}$}{$\found \gets \texttt{true}$}
      $e \gets \pi(e)$\;
    }
    \lIf{$e = d$}{\Return{\found}}
    \Return{\normalfont \texttt{null}}\;
  }
\end{algorithm}

>>>>

\newcommand\pfun{\mathrel{\ooalign{\hfil$\mapstochar\mkern5mu$\hfil\cr$\to$\cr}}}
\newcommand\mdoubleplus{\mathbin{+\mkern-10mu+}}

@@

\paragraph{Things I might need to explain.}
\begin{itemize}
\item atom, (positive/negative) literal, constant, predicate, variable, literal variable, clause, unit clause
\item $\Vars$, $\Vars(c) = \Vars(L) \cup \Vars(C)$
\item $\Doms$ on both formulas and clauses. $\Doms(c) = \Imm \delta_c$, and $\Doms(\phi) = \bigcup_{c \in \phi} \Doms(c)$.
\item maybe: notation for partial function, notation for powerset, notation for disjoint union
\item $\WMC$, $\wwp$, $\wwn$, $\Imm$
\item two parts: compilation and inference.
\item introduce and use arrows for bijections, injections, set inclusions, etc.
\item $\emptyset$ for an empty partial map.
\item notation for variable substitution for literal and constraint sets: $S[x/V]$, where $S$ is the set, $V$ is a set of variable, and $x$ is the new constant or variable. Note that for constraint sets this could mean that a variable and a constant need to switch places.
\item During inference, there is a domain size map $\size\colon \mathcal{D} \to \mathbb{N}_0$.
\item $\sqcup$ for both sets and functions. TODO: I think I use $\cup$ elsewhere.
\item mention which rules are in $\Gamma$ and which ones are in $\Delta$ (and why tryCache has to be in $\Delta$).
\end{itemize}

\paragraph{TODO}
\begin{itemize}
\item should I always say `compilation rule' instead of `operator'?
\item maybe $\pi$ is global enough to have a more unique name
\item capitalise variable names.
\item maybe introduce/use notation for vertices, edges of a graph
\item maybe put each function into its own algorithm environment, using the caption for the function name
\end{itemize}

@@

\begin{definition}
  A \emph{domain} is a symbol for a finite set.\footnote{In the context of functions, the domain of a function $f$ retains its usual meaning and is denoted $\dom(f)$.} Let $\mathcal{D}$ be the set of all domains and $\mathcal{C} \subset \mathcal{D}$ be the subset of domains introduced as a consequence of constraint removal. Note that both sets (can) expand during the compilation phase.

Let $\pi\colon \mathcal{D} \pfun \mathcal{D}$ be a partial endomorphism on $\mathcal{D}$ that denotes the \emph{parent} relation, i.e., if $\pi(d) = e$ for some $d, e \in \mathcal{D}$, then we call $e$ the parent (domain) of $d$, and $e$ a child of $d$. Intuitively, $\pi$ arranges all domains into a forest---thus, we often use graph theoretical terminology to describe properties of and relationships between domains.
\end{definition}

@@

\begin{definition}
  A \emph{clause} is a triple $c = (L, C, \delta_c)$, where $L$ is the set of literals, $C$ is a set of inequality constraints, and $\delta_c\colon \Vars(c) \to \mathcal{D}$ is a function that maps all variables in $c$ to their domains such that if $(x, y) \in C$ for some variables $x$ and $y$, then $\delta_c(x) = \delta_c(y)$. Equality of clauses is defined in the usual way (i.e., all variables, domains, etc. must match). TODO: we will always use this subscript notation for the $\delta$'s. 
\end{definition}

@@

TODO: I'm already using the square brackets to denote lists.

@@

\begin{definition}
  A \emph{first-order deterministic decomposable negation normal form computational graph} (FCG) is a (weakly connected) directed graph with a a single source, node labels, and ordered direct successors. We denote an FCG as $G = (V, s, N^+, \lambda)$, where $V$ is the set of nodes, $s \in V$ is the source, $N^+$ is the direct successor function that maps each node in $V$ to a list that contains either other nodes in $V$ or $\star$, which means that the target of the edge is yet to be determined, and $\lambda\colon V \to \mathcal{L}$ is the vertex-labelling function with some codomain $\mathcal{L}$, which is assumed to be the same for all FCGs. For each $v \in V$, the length of the list $N^+(v)$ is determined by $\lambda(v)$.

@@

\paragraph{TODO.}
\begin{itemize}
\item explain the connection to the algebraic formalism (or just don't use it). % We will also use algebraic notation to describe (parts of) circuits: circuittype_parameters(nodes pointed to by the outedges)
\item define `an FCG for a formula'
\item vertex or node, $v$ or $n$?
\item what notation for lists (a.k.a. sequences) am I going to use? $\mdoubleplus$ for concatenation (of both two lists and an element to a list), $\in$ for (in-order) enumeration, $[]$ for an empty list, $[x]$ for a list with one element, $|L|$ for the length of the list, and $h : t$ to denote a list with first element (a.k.a. head) $h$ and remaining list (a.k.a. tail) $t$. Also add an example of the list comprehension notation.
\end{itemize}

@@

\begin{algorithm}[t]
  \caption{The (main part of the) search algorithm}
  \label{alg:search}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKwData{solutions}{solutions}
  \SetKwFunction{applyGreedyRules}{applyGreedyRules}
  \SetKwFunction{applyAllRules}{applyAllRules}
  \SetKwFunction{put}{put}
  \SetKwFunction{get}{get}
  \SetKwFunction{emptyy}{empty}
  \KwIn{a formula $\phi_0$}
  \KwResult{all found FCGs for $\phi_0$ are in the set \solutions}
  $\solutions \gets \emptyset$\;
  $C_0 \gets \emptyset$\;
  $(G_0, C_0, L_0) \gets \applyGreedyRules{$\phi_0$, $C_0$}$\;
  \lIf{$L_0 = []$}{$\solutions \gets \{\, G_0 \,\}$}
  \Else{
    $q \gets \text{an empty queue of states}$\;
    $q.\put{$(G_0, C_0, L_0)$}$\;
    \While{{\bf not} $q.\emptyy{}$}{
      \ForEach{state $(G, C, L) \in \applyAllRules{$q.\get{}$}$}{
        \lIf{$L = []$}{$\solutions \gets \solutions \cup \{\, G \,\}$}
        \lElse{$q.\put{$(G, C, L)$}$}
      }
    }
  }
\end{algorithm}

\begin{algorithm}
  \caption{Functions used in \cref{alg:search} for applying compilation rules}
  \label{alg:apply}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKwData{newStates}{newStates}
  \SetKwFunction{applyGreedyRules}{applyGreedyRules}
  \SetKwFunction{applyAllRules}{applyAllRules}
  \SetKwFunction{applyGreedyRulesToAllFormulas}{applyGreedyRulesToAllFormulas}
  \SetKwFunction{mergeFcgs}{mergeFcgs}
  \SetKwFunction{updateCache}{updateCache}
  \KwData{a set of greedy rules $\Gamma$}
  \KwData{a set of non-greedy rules $\Delta$}
  \Fn{\applyGreedyRules{$\phi$, $C$}}{
    \ForEach{rule $r \in \Gamma$}{
      $S \gets r(\phi)$\;
      \If{$S \ne \emptyset$}{
        $(G, L) \gets \text{any element of } S$\;
        \lIf{$G = {\normalfont \texttt{null}}$}{\Return{\applyGreedyRules{the only element of $L$, $C$}}}
        \Else{
          $(V, s, N^+, \lambda) \gets G$\;
          $C \gets \updateCache(C, \phi, s)$\;
          \Return{\applyGreedyRulesToAllFormulas{$G$, $C$, $L$}}\;
        }
      }
    }
    \Return{$({\normalfont \texttt{null}}, C, [\phi])$}\;
  }
  \Fn{\applyGreedyRulesToAllFormulas{$G$, $C$, $L$}}{
    $(V, s, N^+, \lambda) \gets G$\;
    $N^+(s) \gets []$\;
    $L' \gets []$\;
    \ForEach{formula $\phi \in L$}{
      $(G', C, L'') \gets \applyGreedyRules{$\phi$, $C$}$\;
      $L' \gets L' \mdoubleplus L''$\;
      \lIf{$G' = {\normalfont \texttt{null}}$}{$N^+(s) \gets N^+(s) \mdoubleplus [\star]$}
      \Else{
        $(V', s', N', \lambda') \gets G'$\;
        $V \gets V \sqcup V'$\;
        $N^+ \gets N^+ \sqcup N'$\;
        $N^+(s) \gets N^+(s) \mdoubleplus [s']$\;
        $\lambda \gets \lambda \sqcup \lambda'$\;
      }
    }
    \Return{$((V, s, N^+, \lambda), C, L')$}\;
  }
  \Fn{\applyAllRules{$s$}}{
    $(G, C, L) \gets s$\;
    $\phi : T \gets L$\;
    $(G', C', L') \gets \text{a copy of } s$\;
    $\newStates \gets []$\;
    \ForEach{rule $r \in \Delta$}{
      \ForEach{$(G'', L'') \in r(\phi)$}{
        \lIf{$G'' = {\normalfont \texttt{null}}$}{$\newStates \gets \newStates \mdoubleplus \applyAllRules{$(G', C', L'')$}$}
        \Else{
          $(V, s, N^+, \lambda) \gets G''$\;
          $C' \gets \updateCache{$C'$, $\phi$, $s$}$\;
          $(G'', C', L'') \gets \applyGreedyRulesToAllFormulas{$G''$, $C'$, $L''$}$\;
          \lIf{$G' = {\normalfont \texttt{null}}$}{$\newStates \gets \newStates \mdoubleplus [(G'', C', L'' \mdoubleplus T)]$}
          \lElse{$\newStates \gets \newStates \mdoubleplus [(\mergeFcgs{$G'$, $G''$}, C', L'' \mdoubleplus T)]$}
        }
      }
      $(G', C', L') \gets \text{a copy of } s$\;
    }
    \Return{\newStates}\;
  }
\end{algorithm}

\begin{algorithm}
  \caption{Helper functions used by \cref{alg:apply}}
  \label{alg:helpers}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKwFunction{mergeFcgs}{mergeFcgs}
  \SetKwFunction{updateCache}{updateCache}
  \Fn{\updateCache{$C$, $\phi$, $v$}}{
    \lIf{$\#\phi \not\in \dom(C)$}{\Return{$C \cup \{\, \#\phi \mapsto (\phi, v) \,\}$}}
    \lIf{there is no $(\phi', v') \in C(\#\phi)$ such that $v' = v$}{$C(\#\phi) \gets (\phi, v) \mdoubleplus C(\#\phi)$}
    \Return{$C$}\;
  }
  \Fn{\mergeFcgs{$G = (V, s, N^+, \lambda)$, $G' = (V', s', N', \lambda')$, $r = s$}}{
    \lIf{$\lambda(r) = \textsc{Ref}$}{\Return{\normalfont \texttt{null}}}
    \ForEach{$t \in N^+(r)$}{
      \If{$t = \star$}{
        replace $t$ with $s'$ in $N^+(r)$\;
        \Return{$(V \sqcup V', s, N^+ \sqcup N', \lambda \sqcup \lambda')$}\;
      }
      $G'' \gets \mergeFcgs{$G$, $G'$, $t$}$\;
      \lIf{$G'' \ne {\normalfont \texttt{null}}$}{\Return{$G''$}}
    }
    \Return{\normalfont \texttt{null}}\;
  }
\end{algorithm}

TODO: explain the `tail' part of the algorithm, i.e., that the first formula is replaced by some nodes and some formulas.

@@

\begin{algorithm}
  \SetKwData{changed}{changed}
  \caption{Propagating atoms for smoothing across the FCG in a way that avoids infinite loops}
  \label{alg:smoothing}
  \KwIn{FCG $(V, s, N^+, \lambda)$}
  \KwIn{function $\iota$ that maps labels in $\mathcal{L}$ to sets of atoms}
  \KwIn{functions $\{\,f_l\,\}_{l \in \mathcal{L}}$ that take a list of sets of atoms and return a set of atoms}
  \KwOut{function $S$ that maps vertices in $V$ to sets of atoms}
  $S \gets \{\, v \mapsto \iota(\lambda(v)) \mid v \in V \,\}$\;
  $\changed \gets \texttt{true}$\;
  \While{\changed}{
    $\changed \gets \texttt{false}$\;
    \ForEach{vertex $v \in V$}{
      $S' \gets f_{\lambda(v)}([S(w) \mid w \in N^+(v)])$\;
      \If{$S' \ne S(v)$}{
        $\changed \gets \texttt{true}$\;
        $S(v) \gets S'$\;
      }
    }
  }
\end{algorithm}

@@

\begin{definition}[Notation]
  For any clause $c = (L, C, \delta_c)$, bijection $\beta\colon \Vars(c) \to V$ (for some set of variables $V$) and function $\gamma\colon \Doms(c) \to \mathcal{D}$, let $c[\beta, \gamma] = d$ be the clause with all occurrences of any variable $v \in \Vars(c)$ in $L$ and $C$ replaced with $\beta(v)$ (so $\Vars(d) = V$) and $\delta_d\colon V \to \mathcal{D}$ defined as $\delta_d \coloneqq \gamma \circ \delta_c \circ \beta^{-1}$. In other words, $\delta_d$ is the unique function that makes the following diagram commute:
  \[
  \begin{tikzcd}
    \Vars(c) \ar[r, tail, two heads, "\beta"] \arrow[d, swap, "\delta_c"] & V = \Vars(d) \ar[d, dashed, "\exists!\delta_d"] \\
    \Doms(c) \ar[r, swap, "\gamma"] & \mathcal{D}.
  \end{tikzcd}
  \]
\end{definition}

\begin{algorithm}[t]
  \caption{A recursive function for checking whether one can reuse the circuit for computing $\WMC(\phi)$ to compute $\WMC(\psi)$. Both $\phi$ and $\psi$ are formulas, and $\rho\colon \Doms(\phi) \protect\pfun \Doms(\psi)$ is a partial map.}
  \label{alg:recursion}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKw{KwRet}{yield}
  \SetKwFunction{identifyRecursion}{identifyRecursion}
  \SetKwFunction{findHistory}{traceAncestors}
  \SetKwFunction{generateMaps}{generateMaps}
  \SetKwFunction{constructDomainMap}{constructDomainMap}
  \SetKwData{suitable}{suitableBijection}
  \SetKwData{found}{foundConstraintRemoval}
  \Fn{\identifyRecursion{$\phi$, $\psi$, $\rho = \emptyset$, $\found = {\normalfont \texttt{false}}$}}{
    \lIf{$|\phi| \ne |\psi|$ {\bf or} $\#\phi \ne \#\psi$}{\Return{\normalfont \texttt{null}}}
    \If{$\phi = \emptyset$}{
      \lIf{$\found$}{\Return{$\rho$}}
      \Return{\normalfont \texttt{null}}\;
    }
    \ForEach{clause $c \in \phi$\label{line:for1}}{
      \ForEach{clause $d \in \psi$ such that $\#d = \#c$\label{line:for2}} {
        \ForEach{$(\beta, \gamma) \in \generateMaps{$c$, $d$, $\rho$}$ such that $c[\beta, \gamma] = d$\label{line:generateMaps}}{
          $\found' \gets \found$\;
          $\suitable \gets \texttt{true}$\;
          \ForEach{$e \in \Doms(c)$\label{line:vars}}{
            $\found'' \gets \findHistory{$e$, $\gamma(e)$}$\;
            \If{$\found'' = {\normalfont \texttt{null}}$}{
              $\suitable \gets \texttt{false}$\;
              break\;
            }
            \lIf{$\found''$}{$\found' \gets \texttt{true}$}
          }
          \If{\suitable}{
            $\rho'' \gets \identifyRecursion{$\phi \setminus \{\, c \,\}$, $\psi \setminus \{\, d \,\}$, $\rho \cup \gamma$, $\found'$}$\;\label{line:recursion}
            \lIf{$\rho'' \ne {\normalfont \texttt{null}}$}{\Return{$\rho''$}}
          }
        }
      }
      \Return{\normalfont \texttt{null}}\;
    }
  }
  \Fn{\generateMaps{$c$, $d$, $\rho$}}{
    \ForEach{bijection $\beta\colon \Vars(c) \to \Vars(d)$\label{line:bijection}}{
      $\gamma \gets \constructDomainMap{$\Vars(c)$, $\delta_c$, $\delta_d$, $\beta$, $\rho$}$\;
      \lIf{$\gamma \ne {\normalfont \texttt{null}}$}{\KwRet{$(\beta, \gamma)$}}
    }
  }
  \Fn{\constructDomainMap{$V$, $\delta_c$, $\delta_d$, $\beta$, $\rho$}}{
    $\gamma \gets \emptyset$\;
    \ForEach{$v \in V$}{
      \lIf{$\delta_c(v) \in \dom(\rho)$ {\bf and} $\rho(\delta_c(v)) \ne \delta_d(\beta(v))$}{\Return{\normalfont \texttt{null}}}
      \lIf{$\delta_c(v) \not\in \dom(\gamma)$}{$\gamma \gets \gamma \cup \{\, \delta_c(v) \mapsto \delta_d(\beta(v)) \,\}$}
      \lElseIf{$\gamma(\delta_c(v)) \ne \delta_d(\beta(v))$}{\Return{\normalfont \texttt{null}}}
    }
    \Return{$\gamma$}\;
  }
  \Fn{\findHistory{$d$, $e$}}{
    $\found \gets \texttt{false}$\;
    \While{$e \ne d$ {\bf and} $e \in {\normalfont \dom}(\pi)$}{
      \lIf{$e \in \mathcal{C}$}{$\found \gets \texttt{true}$}
      $e \gets \pi(e)$\;
    }
    \lIf{$e = d$}{\Return{\found}}
    \Return{\normalfont \texttt{null}}\;
  }
\end{algorithm}
